Fast and scalable Gaussian process modeling with applications to astronomical time series \vspace{-3\baselineskip}

### 2017-03-15 paper/ms.tex
summary of benchmark/scaling results

The benchmark experiments shown here were performed on a MacBook Pro with two
2.6~GHz CPUs but we find similar results on a Dell workstation with 16 2.7~GHz
CPUs and running Ubuntu.

NNZ: $(10\,J + 1)\,N - 4\,J + 1$

\begin{figure}[!htbp]
\begin{center}
\includegraphics[width=\textwidth]{../examples/benchmark/benchmark_darwin.pdf}
\caption{A benchmark showing the computational scaling of \celerite\ with a
    vanilla band solver as a function of the number of data points and the
    number of terms.
    \emph{(left)} The cost of computing \eq{gp-likelihood} with a covariance
    matrix given by \eq{celerite-kernel} as a function of the number of data
    points $N$.
    The different colors show the cost for different numbers of terms $J$ as
    listed in the legend.
    To guide the eye, the black line shows linear scaling in $N$.
    \emph{(right)} The same information plotted as a function of $J$ for
    different values of $N$.
    The legend shows the value of $N$ for each color and the black line shows
    quadratic scaling in $J$.
    \figurelabel{benchmark}}
\end{center}
\end{figure}

\begin{figure}[!htbp]
\begin{center}
\includegraphics[width=\textwidth]{../examples/benchmark/benchmark_darwin_lapack.pdf}
\caption{The same as \Figure{benchmark} but using the band solver from
    \project{LAPACK} with the \project{MKL} \project{BLAS} library.
    \figurelabel{benchmark-lapack}}
\end{center}
\end{figure}

\begin{figure}[!htbp]
\begin{center}
\includegraphics[width=\textwidth]{../examples/benchmark/benchmark_darwin_sparse.pdf}
\caption{The same as \Figure{benchmark} but using the \texttt{SparseLU} solver
    from \project{Eigen} \citep{Guennebaud:2010}.
    \figurelabel{benchmark-sparse}}
\end{center}
\end{figure}

>>>>

The benchmark experiments shown here were performed on a MacBook Pro with two
2.6~GHz CPUs but we find similar results on a Dell workstation with 16 2.7~GHz
CPUs and running Ubuntu.
\Figure{benchmark} shows how the cost of computing \eq{gp-likelihood} scales
with $N$ and $J$ using the \texttt{vanilla} solver.
As expected theoretically, the scaling is linear in $N$ for all $N$ and cubic
in $J$ for large $J$.
\Figure{benchmark-lapack} and \Figure{benchmark-sparse} are the same plots for
the \texttt{lapack} and \texttt{sparse} solvers respectively.
For each of these optimized solvers, the empirical scaling is
$\mathcal{O}(N\,J^2)$ at the cost of some extra overhead.
Therefore, for $J \lesssim 5$ or $10$ the \texttt{vanilla} solver is more
efficient than the other algorithms.
The real world performance of \celerite\ depends on the specific platform,
hardware, and \project{LAPACK}/\project{BLAS} implementation but we have found
qualitatively similar results across popular platforms and state-of-the-art
libraries.

\begin{figure}[!htbp]
\begin{center}
\includegraphics[width=\textwidth]{../examples/benchmark/benchmark_darwin.pdf}
\caption{A benchmark showing the computational scaling of \celerite\ using the
    \texttt{vanilla} band solver as a function of the number of data points
    and the number of terms.
    \emph{(left)} The cost of computing \eq{gp-likelihood} with a covariance
    matrix given by \eq{celerite-kernel} as a function of the number of data
    points $N$.
    The different colors show the cost for different numbers of terms $J$ as
    listed in the legend.
    To guide the eye, the black line shows linear scaling in $N$.
    \emph{(right)} The same information plotted as a function of $J$ for
    different values of $N$.
    The legend shows the value of $N$ for each color and the black line shows
    quadratic scaling in $J$.
    \figurelabel{benchmark}}
\end{center}
\end{figure}

\begin{figure}[!htbp]
\begin{center}
\includegraphics[width=\textwidth]{../examples/benchmark/benchmark_darwin_lapack.pdf}
\caption{The same as \Figure{benchmark} but using the \texttt{lapack} solver
    with the \project{MKL} \project{BLAS} library.
    \figurelabel{benchmark-lapack}}
\end{center}
\end{figure}

\begin{figure}[!htbp]
\begin{center}
\includegraphics[width=\textwidth]{../examples/benchmark/benchmark_darwin_sparse.pdf}
\caption{The same as \Figure{benchmark} but using the \texttt{sparse} solver
    from \project{Eigen} \citep{Guennebaud:2010}.
    \figurelabel{benchmark-sparse}}
\end{center}
\end{figure}
